### 2023-04-05 outputs/paper/paper.qmd
General updates
- Added citations to paper
- Renamed documents
- Fixed year parsing error in data cleaning

---
title: "Bacon, Eggs, and Inflation: Can household groceries track inflation?"
format: pdf
author: "Sebastian Rodriguez"
abstract: "Canada’s average inflation rate in 2022 was 6.7%, a figure not seen since the 1980s. This inflation rate reflects an increase in the cost of goods and services (determined by the consumer price index) and impacts the wallets of people across the nation. Studies show that the price of household groceries directly influences inflation expectations. This paper investigates whether the cost of groceries can track Canada’s inflation rate, examining common items like bacon, eggs, chicken, milk, frozen peas, and pasta sauce. Preliminary data from Statistics Canada and the Bank of Canada reveals that the cost rate of common groceries often exceeds the rate of inflation, and Canadians should not solely rely on grocery prices to assess the inflation rate."
date: today
date-format: long
editor: visual
toc: TRUE
toc-depth: 1
number-sections: TRUE
bibliography: references.bib
thanks: "See the GitHub repository supporting this analysis: https://github.com/seb646/groceries-and-inflation"
---
```{r}
#| messages: false
#| echo: false
#| warning: false
#| output: false
#### Preamble ####
# Title: Bacon, Eggs, and Inflation: Can household groceries track inflation?
# Author: Sebastian Rodriguez
# Email: me@srod.ca
# Date: TBD
# GitHub: ttps://github.com/seb646/groceries-and-inflation

#### Workspace set-up ####
# Define required libraries 
library(lubridate)
library(tidyverse)
library(dplyr)
library(knitr)
library(janitor)
library(scales)
library(RColorBrewer)
library(ggplot2)
library(kableExtra)
library(here)
library(readr)
library(arrow)
library(gridExtra)

# Fetch the the clean response data
inflation_data <- arrow::read_parquet(here::here("inputs/data/clean_inflation.parquet"), show_col_types = FALSE)

# Fetch clean performance data
grocery_data <- arrow::read_parquet(here::here("inputs/data/clean_grocery.parquet"), show_col_types = FALSE)

inflation_graph <- inflation_data |>
  # Plot graphs, define x axis
  ggplot(aes(x = year, y = inflation_data$inflation_rate, group = 1)) +

  # Define axis labels
  labs(
    y = "Inflation Rate (%)",
    x = "Year"
  ) +

  # Set the theme
  theme_minimal() +

  theme(
    axis.title.x = element_text(size = 15),
    axis.text.x = element_text(size = 14, angle = 50, vjust = 0.5),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 14),
  ) +

  # Define styles for axis, legends, and labels
  theme(
    axis.text.x = element_text(angle = 50, vjust = 0.5),
  ) +

  # Render line
  geom_line()
```

## Introduction

Inflation is the pervasive and sustained increase in the cost of goods and services, and it has become a significant concern for economies worldwide. Inflation is most often observed as an impact on a currency's purchasing power, but it also has far-reaching effects on the affordability of household goods. In Canada, the inflation rate is tracked by the consumer price index, or CPI, which is an amalgamation of the cost of goods and services around the country [@cite_boc_inflation]. For example, if the price of energy increases, so too does the CPI, which in turn increases the rate of inflation. If the inflation rate becomes too high (or too low), it can indicate serious problems for a country's economy and has lasting financial implications for the country's population.

In 2022, Canada's average rate of inflation was 6.7% -- a record high not seen since the 1980s [@cite_td]. This reflected a rise in the price of goods and services, like energy and food, for Canadians. One of the areas where inflation rates are arguably most observable is daily grocery store purchases. In fact, one study shows that "household-level grocery price changes significantly shape\[s\] inflation expectations" [@cite_study_1]. It is important to note that the researchers specify expectations, rather than reality, as inflation is tracked by the larger CPI, which includes housing, energy, clothing, and other items in addition to food. Still, this perception influences how Canadians understand their country's economic situation. As food is a factor considered by the CPI, it is logical to draw a connection between the two factors.

This paper investigates whether grocery store items can track Canada's inflation rate. It uses data from Statistics Canada to measure the 5-year cost history of common grocery store items, including bacon, eggs, chicken, milk, frozen peas, and pasta sauce [@cite_data_statscan]. It also uses data from the Bank of Canada to graph the yearly averages of inflation rates from 2017 to 2022 [@cite_data_boc]. By examining the percent increases in the cost of these food items, it explores the existence of a relationship between the cost of food and inflation to determine whether food is a useful indicator in tracking inflation.

Preliminary results indicate that food is not a reliable metric for tracking inflation, as the cost of food increased at rates beyond inflation during the observed period. While food cost certainly plays a role in the country's overall CPI and inflation rate. Canadians should not solely depend on the cost of food to assess the country's inflation rate.

\newpage

## Measurement

## Inflation Rates and Grocery Products

### Bacon (per 500 grams)

```{r}
#| messages: false
#| echo: false
#| warning: false
#| label: fig-3
#| fig-cap: "Inflation Rate vs. Cost of Bacon (since 2017)"
#| fig-align: center
#| fig-width: 10
#| fig-height: 3
#### Create graphs plotting mandated and actual response times ####
bacon <- grocery_data |>
  # Plot graphs, define x axis
  ggplot(aes(x = year, y = grocery_data$bacon, group = 1)) +

  # Define axis labels
  labs(
    y = "Cost ($)",
    x = "Year"
  ) +

  # Set the theme
  theme_minimal() +

  # Define styles for axis and labels
  theme(
    axis.title.x = element_text(size = 15),
    axis.text.x = element_text(size = 14, angle = 50, vjust = 0.5),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 14),
  ) +

  scale_y_continuous(limits=c(5, 10)) +

  # Render lines
  geom_line()

grid.arrange(bacon, inflation_graph, ncol=2)
```

```{r}
#| messages: false
#| echo: false
#| warning: false
#| label: fig-4
#| fig-cap: "Bacon Cost Rate and Inflation (since 2017)"
#| fig-align: center
#| fig-width: 10
#| fig-height: 3
#### Create graphs plotting mandated and actual response times ####
# Define colors to be used in legend
bacon_table <- grocery_data |>
  select(year, bacon)

bacon_inflation_data <-
  # Specify which tables to merge (this merges by row)
  merge(
    x = bacon_table,
    y = inflation_data, 
    by = "year",
    all.x = TRUE
  ) |>

  # Calculate the percent change of chicken prices
  mutate(percent_change = round((((bacon - lag(bacon, 1)) / lag(bacon, 1)) * 100), digits = 2)) |>

  # Remove empty rows
  drop_na() |>

  # Change months to fiscal quarters
  mutate(bacon =  paste("\\$", bacon, sep = "")) |>
  mutate(percent_change =  paste(percent_change, "\\%", sep = "")) |>
  mutate(inflation_rate =  paste(inflation_rate, "\\%", sep = "")) |>

  # Order the columns 
  select(year, bacon, percent_change, inflation_rate) |>

  # Rename the columns
  rename(
    "Year" = "year", 
    "Bacon Cost" = "bacon", 
    "Inflation Rate" = "inflation_rate", 
    "Bacon Cost Rate" = "percent_change"
  )

bacon_inflation_data |>
  knitr::kable(align = "l", booktabs = TRUE, escape = FALSE, linesep = "" )|>
  kable_styling(font_size = 10, latex_options = "hold_position", full_width = TRUE)

```

\newpage

### Eggs (per 1 dozen)

```{r}
#| messages: false
#| echo: false
#| warning: false
#| label: fig-7
#| fig-cap: "Inflation Rate vs. Cost of Eggs (since 2017)"
#| fig-align: center
#| fig-width: 10
#| fig-height: 3
#### Create graphs plotting mandated and actual response times ####
eggs <- grocery_data |>
  # Plot graphs, define x axis
  ggplot(aes(x = year, y = grocery_data$eggs, group = 1)) +

  # Define axis labels
  labs(
    y = "Cost ($)",
    x = "Year"
  ) +

  # Set the theme
  theme_minimal() +

  # Define styles for axis and labels
  theme(
    axis.title.x = element_text(size = 15),
    axis.text.x = element_text(size = 14, angle = 50, vjust = 0.5),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 14),
  ) +

  scale_y_continuous(limits=c(0, 5)) +

  # Render lines
  geom_line()

grid.arrange(eggs, inflation_graph, ncol=2)
```

```{r}
#| messages: false
#| echo: false
#| warning: false
#| label: fig-8
#| fig-cap: "Eggs Cost Rate and Inflation (since 2017)"
#| fig-align: center
#| fig-width: 10
#| fig-height: 3
#### Create graphs plotting mandated and actual response times ####
# Define colors to be used in legend
eggs_table <- grocery_data |>
  select(year, eggs)

eggs_inflation_data <-
  # Specify which tables to merge (this merges by row)
  merge(
    x = eggs_table,
    y = inflation_data, 
    by = "year",
    all.x = TRUE
  ) |>

  # Calculate the percent change of chicken prices
  mutate(percent_change = round((((eggs - lag(eggs, 1)) / lag(eggs, 1)) * 100), digits = 2)) |>

  # Remove empty rows
  drop_na() |>

  # Change months to fiscal quarters
  mutate(eggs =  paste("\\$", eggs, sep = "")) |>
  mutate(percent_change =  paste(percent_change, "\\%", sep = "")) |>
  mutate(inflation_rate =  paste(inflation_rate, "\\%", sep = "")) |>

  # Order the columns 
  select(year, eggs, percent_change, inflation_rate) |>

  # Rename the columns
  rename(
    "Year" = "year", 
    "Eggs Cost" = "eggs", 
    "Inflation Rate" = "inflation_rate", 
    "Eggs Cost Rate" = "percent_change"
  )

eggs_inflation_data |>
  knitr::kable(align = "l", booktabs = TRUE, escape = FALSE, linesep = "" )|>
  kable_styling(font_size = 10, latex_options = "hold_position", full_width = TRUE)

```

\newpage

### Chicken Breasts (per 3 kilograms)

```{r}
#| messages: false
#| echo: false
#| warning: false
#| label: fig-1
#| fig-cap: "Inflation Rate vs. Cost of Chicken Breasts (since 2017)"
#| fig-align: center
#| fig-width: 10
#| fig-height: 3
#### Create graphs plotting mandated and actual response times ####
# Define colors to be used in legend
chicken_graph <- grocery_data |>
  # Plot graphs, define x axis
  ggplot(aes(x = year, y = grocery_data$chicken, group = 1)) +

  # Define axis labels
  labs(
    y = "Cost ($)",
    x = "Year"
  ) +

  # Set the theme
  theme_minimal() +

  # Define styles for axis and labels
  theme(
    axis.title.x = element_text(size = 15),
    axis.text.x = element_text(size = 14, angle = 50, vjust = 0.5),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 14),
  ) +

  scale_y_continuous(limits=c(10, 15)) +

  # Render lines
  geom_line()

grid.arrange(chicken_graph, inflation_graph, ncol=2)
```

```{r}
#| messages: false
#| echo: false
#| warning: false
#| label: fig-2
#| fig-cap: "Price of Chicken Breasts Cost Rate and Inflation (since 2017)"
#| fig-align: center
#| fig-width: 10
#| fig-height: 3
#### Create graphs plotting mandated and actual response times ####
# Define colors to be used in legend
chicken_table <- grocery_data |>
  select(year, chicken)

chicken_inflation_data <-
  # Specify which tables to merge (this merges by row)
  merge(
    x = chicken_table,
    y = inflation_data, 
    by = "year",
    all.x = TRUE
  ) |>

  # Calculate the percent change of chicken prices
  mutate(percent_change = round((((chicken - lag(chicken, 1)) / lag(chicken, 1)) * 100), digits = 2)) |>

  # Remove empty rows
  drop_na() |>

  # Change months to fiscal quarters
  mutate(chicken =  paste("\\$", chicken, sep = "")) |>
  mutate(percent_change =  paste(percent_change, "\\%", sep = "")) |>
  mutate(inflation_rate =  paste(inflation_rate, "\\%", sep = "")) |>

  # Order the columns 
  select(year, chicken, percent_change, inflation_rate) |>

  # Rename the columns
  rename(
    "Year" = "year",
    "Chicken Cost" = "chicken",
    "Inflation Rate" = "inflation_rate",
    "Chicken Cost Rate" = "percent_change"
  )

# summary <- chicken_inflation_data |>
#     summarize(across(where(is.numeric), mean, na.rm = TRUE))
# 
# chicken_inflation_data <- bind_rows(
#   chicken_inflation_data,
#   summary
# )
# summary

chicken_inflation_data |>
  knitr::kable(align = "l", booktabs = TRUE, escape = FALSE, linesep = "" )|>
  kable_styling(font_size = 10, latex_options = "hold_position", full_width = TRUE)
  # pack_rows("Average", start_row = 6, end_row = 6)

```

\newpage

### Milk (per 2 litres)

```{r}
#| messages: false
#| echo: false
#| warning: false
#| label: fig-5
#| fig-cap: "Inflation Rate vs. Cost of Milk (since 2017)"
#| fig-align: center
#| fig-width: 10
#| fig-height: 3
#### Create graphs plotting mandated and actual response times ####
milk <- grocery_data |>
  # Plot graphs, define x axis
  ggplot(aes(x = year, y = grocery_data$milk, group = 1)) +

  # Define axis labels
  labs(
    y = "Cost ($)",
    x = "Year"
  ) +

  # Set the theme
  theme_minimal() +

  # Define styles for axis and labels
  theme(
    axis.title.x = element_text(size = 15),
    axis.text.x = element_text(size = 14, angle = 50, vjust = 0.5),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 14),
  ) +

  scale_y_continuous(limits=c(0, 5)) +

  # Render lines
  geom_line()

grid.arrange(milk, inflation_graph, ncol=2)
```

```{r}
#| messages: false
#| echo: false
#| warning: false
#| label: fig-6
#| fig-cap: "Milk Cost Rate and Inflation (since 2017)"
#| fig-align: center
#| fig-width: 10
#| fig-height: 3
#### Create graphs plotting mandated and actual response times ####
# Define colors to be used in legend
milk_table <- grocery_data |>
  select(year, milk)

milk_inflation_data <-
  # Specify which tables to merge (this merges by row)
  merge(
    x = milk_table,
    y = inflation_data, 
    by = "year",
    all.x = TRUE
  ) |>

  # Calculate the percent change of chicken prices
  mutate(percent_change = round((((milk - lag(milk, 1)) / lag(milk, 1)) * 100), digits = 2)) |>

  # Remove empty rows
  drop_na() |>

  # Change months to fiscal quarters
  mutate(milk =  paste("\\$", milk, sep = "")) |>
  mutate(percent_change =  paste(percent_change, "\\%", sep = "")) |>
  mutate(inflation_rate =  paste(inflation_rate, "\\%", sep = "")) |>

  # Order the columns 
  select(year, milk, percent_change, inflation_rate) |>

  # Rename the columns
  rename(
    "Year" = "year", 
    "Milk Cost" = "milk", 
    "Inflation Rate" = "inflation_rate", 
    "Milk Cost Rate" = "percent_change"
  )

milk_inflation_data |>
  knitr::kable(align = "l", booktabs = TRUE, escape = FALSE, linesep = "" )|>
  kable_styling(font_size = 10, latex_options = "hold_position", full_width = TRUE)

```

\newpage

### Frozen Peas (per 750 grams)

```{r}
#| messages: false
#| echo: false
#| warning: false
#| label: fig-9
#| fig-cap: "Frozen Peas Cost Rate and Inflation (since 2017)"
#| fig-align: center
#| fig-width: 10
#| fig-height: 3
#### Create graphs plotting mandated and actual response times ####
frozen_peas <- grocery_data |>
  # Plot graphs, define x axis
  ggplot(aes(x = year, y = grocery_data$frozen_peas, group = 1)) +

  # Define axis labels
  labs(
    y = "Cost ($)",
    x = "Year"
  ) +

  # Set the theme
  theme_minimal() +

  # Define styles for axis and labels
  theme(
    axis.title.x = element_text(size = 15),
    axis.text.x = element_text(size = 14, angle = 50, vjust = 0.5),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 14),
  ) +

  scale_y_continuous(limits=c(0, 5)) +

  # Render lines
  geom_line()

grid.arrange(frozen_peas, inflation_graph, ncol=2)
```

```{r}
#| messages: false
#| echo: false
#| warning: false
#| label: fig-10
#| fig-cap: "Inflation Rate vs. Cost of Frozen Peas (since 2017)"
#| fig-align: center
#| fig-width: 10
#| fig-height: 3
#### Create graphs plotting mandated and actual response times ####
# Define colors to be used in legend
frozen_peas_table <- grocery_data |>
  select(year, frozen_peas)

frozen_peas_inflation_data <-
  # Specify which tables to merge (this merges by row)
  merge(
    x = frozen_peas_table,
    y = inflation_data, 
    by = "year",
    all.x = TRUE
  ) |>

  # Calculate the percent change of chicken prices
  mutate(percent_change = round((((frozen_peas - lag(frozen_peas, 1)) / lag(frozen_peas, 1)) * 100), digits = 2)) |>

  # Remove empty rows
  drop_na() |>

  # Change months to fiscal quarters
  mutate(frozen_peas =  paste("\\$", frozen_peas, sep = "")) |>
  mutate(percent_change =  paste(percent_change, "\\%", sep = "")) |>
  mutate(inflation_rate =  paste(inflation_rate, "\\%", sep = "")) |>

  # Order the columns 
  select(year, frozen_peas, percent_change, inflation_rate) |>

  # Rename the columns
  rename(
    "Year" = "year", 
    "Frozen Peas Cost" = "frozen_peas", 
    "Inflation Rate" = "inflation_rate", 
    "Frozen Peas Cost Rate" = "percent_change"
  )

frozen_peas_inflation_data |>
  knitr::kable(align = "l", booktabs = TRUE, escape = FALSE, linesep = "" )|>
  kable_styling(font_size = 10, latex_options = "hold_position", full_width = TRUE)
```

\newpage

### Pasta Sauce (per 650 millilitres)

```{r}
#| messages: false
#| echo: false
#| warning: false
#| label: fig-11
#| fig-cap: "Inflation Rate vs. Cost of Pasta Sauce (since 2017)"
#| fig-align: center
#| fig-width: 10
#| fig-height: 3
#### Create graphs plotting mandated and actual response times ####
pasta_sauce <- grocery_data |>
  # Plot graphs, define x axis
  ggplot(aes(x = year, y = grocery_data$pasta_sauce, group = 1)) +

  # Define axis labels
  labs(
    y = "Cost ($)",
    x = "Year"
  ) +

  # Set the theme
  theme_minimal() +

  # Define styles for axis and labels
  theme(
    axis.title.x = element_text(size = 15),
    axis.text.x = element_text(size = 14, angle = 50, vjust = 0.5),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 14),
  ) +

  scale_y_continuous(limits=c(0, 5)) +

  # Render lines
  geom_line()

grid.arrange(pasta_sauce, inflation_graph, ncol = 2)
```

```{r}
#| messages: false
#| echo: false
#| warning: false
#| label: fig-12
#| fig-cap: "Pasta Sauce Cost Rate and Inflation (since 2017)"
#| fig-align: center
#| fig-width: 10
#| fig-height: 3
#### Create graphs plotting mandated and actual response times ####
# Define colors to be used in legend
pasta_sauce_table <- grocery_data |>
  select(year, pasta_sauce)

pasta_sauce_inflation_data <-
  # Specify which tables to merge (this merges by row)
  merge(
    x = pasta_sauce_table,
    y = inflation_data, 
    by = "year",
    all.x = TRUE
  ) |>

  # Calculate the percent change of chicken prices
  mutate(percent_change = round((((pasta_sauce - lag(pasta_sauce, 1)) / lag(pasta_sauce, 1)) * 100), digits = 2)) |>

  # Remove empty rows
  drop_na() |>

  # Change months to fiscal quarters
  mutate(pasta_sauce =  paste("\\$", pasta_sauce, sep = "")) |>
  mutate(percent_change =  paste(percent_change, "\\%", sep = "")) |>
  mutate(inflation_rate =  paste(inflation_rate, "\\%", sep = "")) |>

  # Order the columns 
  select(year, pasta_sauce, percent_change, inflation_rate) |>

  # Rename the columns
  rename(
    "Year" = "year", 
    "Pasta Sauce Cost" = "pasta_sauce", 
    "Inflation Rate" = "inflation_rate", 
    "Pasta Sauce Cost Rate" = "percent_change"
  )

pasta_sauce_inflation_data |>
  knitr::kable(align = "l", booktabs = TRUE, escape = FALSE, linesep = "" )|>
  kable_styling(font_size = 10, latex_options = "hold_position", full_width = TRUE)
```

\newpage

## Results

## Discussion

\newpage

## Inclusive Acknowledgements

**Land Acknowledgement**

We wish to acknowledge the land on which this data was analyzed. For thousands of years, it has been the traditional land of the Huron-Wendat, the Seneca, and the Mississaugas of the Credit. Today, this meeting place is still the home to many Indigenous people from across Turtle Island and we are grateful to have the opportunity to work on this land.

This statement was originally prepared in consultation with First Nations House and the Elders Circle for use at the University of Toronto [@cite_land_acknowledgement].\

**Data Acknowledgement**

The data used in this paper was obtained through Open Data Toronto, an initiative to provide openly accessible data about the city [@cite_about_opendatatoronto]. This paper uses three data sets from Open Data Toronto, each provided by the Toronto Paramedic Services:

-   Paramedic Services Incident Data[^1]
-   Land Ambulance Response Time Standard[^2]
-   Pre-Hospital Emergency Care Performance Metrics[^3]

[^1]: Paramedic Services Incident Data is available at: https://open.toronto.ca/dataset/paramedic-services-incident-data

[^2]: Land Ambulance Response Time Standard is available at: https://open.toronto.ca/dataset/land-ambulance-response-time-standard

[^3]: Pre-Hospital Emergency Care Performance Metrics is available at: https://open.toronto.ca/dataset/pre-hospital-emergency-care-performance-metrics

**Resources Acknowledgment**

The primary tool used to analyse data in this paper is R, an open-source statistical programming language [@cite_tool_r]. The paper also uses a number of R packages, including: dplyr [@cite_tool_dplyr], ggplot2 [@cite_tool_ggplot2], here [@cite_tool_here], janitor [@cite_tool_janitor], kableExtra [@cite_tool_kableextra], knitr [@cite_tool_knitr], lubridate [@cite_tool_lubridate], opendatatoronto [@cite_tool_opendatatoronto], readr [@cite_tool_readr], RColorBrewer [@cite_tool_rcolorbrewer], scales [@cite_tool_scales], and tidyverse [@cite_tool_tidyverse]. \newpage

## References

<!-- Comment required to show references as a numbered section -->